---
title: "DS105-06-04 - Mixed Measure ANOVA in R Activity"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

## Load Libraries

```{r}
library(tidyverse)
library("rcompanion")
library("car")
```

------------------------------------------------------------------------

## Load in Data

```{r}
breakfast <- read.csv("./assets/breakfast.csv")

  # View data
    head(breakfast)
```

------------------------------------------------------------------------

## Question Setup

`Does eating breakfast in the morning affect weight loss changes from baseline to follow up?`

------------------------------------------------------------------------

## Data Wrangling

Data wrangling for the mixed measures ANOVA is done exactly the same as
it was for the repeated measures ANOVA.

### Filter Data

```{r}
breakfast1 <- breakfast[1:33,]

head(breakfast1)
```

### Reshape Data

```{r}
keep = c("Participant.Code", "Treatment.Group", "Age..y.", "Sex", "Height..m.", "Baseline.Body.Mass..kg.", "Follow.Up.Body.Mass..kg.")

breakfast2 <- breakfast1[keep]

head(breakfast2)
```

------------------------------------------------------------------------

## Baseline Data

```{r}
breakfast3 <- breakfast2[,1:5]
breakfast3$repdat <- breakfast2$Baseline.Body.Mass..kg.
breakfast3$contrasts <- "T1"
head(breakfast3)
```

## Follow-up Data

```{r}
breakfast4 <- breakfast2[,1:5]
breakfast4$repdat <- breakfast2$Follow.Up.Body.Mass..kg.
breakfast4$contrasts <- "T2"
head(breakfast4)
```

Once you have both of those, then you need to `rbind()` them back
together into one whole dataset:

```{r}
breakfast5 <- rbind(breakfast3, breakfast4)
```

------------------------------------------------------------------------

## **Testing Assumptions**

------------------------------------------------------------------------

### Assumption: Normality

```{r}
plotNormalHistogram(breakfast2$Baseline.Body.Mass..kg.)
```

```{r}
plotNormalHistogram(breakfast2$Follow.Up.Body.Mass..kg.)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

### Assumption: Homogeneity of Variance

```{r}
leveneTest(repdat ~ Treatment.Group*contrasts, data=breakfast5)
```

```{r}
str(breakfast5)
```

### **Assumption: Sample Size**

A mixed measures ANOVA requires a sample size of at least 20 per
independent variable or time factor.

We have 1 IV and 1 factor of time = 2 x 20 = need 40 minimum. We have 66
observations.

------------------------------------------------------------------------

## Analysis

You will continue to use the `aov()` function, but add some additional
arguments to it to make it mixed measures.

```{r}
RManova <- aov(repdat~contrasts+Error(Participant.Code), breakfast5)
summary(RManova)
```

::: {style="color: #0187DE"}
# Conclusion:

***`p`*** **`> .5`**, which indicates there is *not a significant
effect* of time on the participants' body mass whether or not they ate
breakfast.
:::

------------------------------------------------------------------------

# `### END`
