---
title: "DS105-03-12 - Analyzing Data in R Hands-On - Bank"
output: html_notebook
---

### Requirements:

> For each part, check and correct for assumptions if possible, perform the appropriate categorical data analysis in R, and provide a one-sentence conclusion at the bottom of your program files about the analysis you performed.

------------------------------------------------------------------------

## Load Libraries

```{r}
install.packages("tidyverse")
library("gmodels")
library("tidyr")
library("tidyverse")
```

## Load Data

```{r}
loans <- read.csv("./assets/loans.csv")

  # VIEW THE DATA
  View(loans)
```

------------------------------------------------------------------------

# Part 1: 

## Question Setup

`Does the term of the loan influence loan status? If so, how?`

## Data Wrangling

The `term` column is formatted as "36 months" but doesn't needed to be recoded because it is a categorical variable.

## Test assumptions and run analysis

```{r}
CrossTable(loans$term, loans$loan_status, fisher=TRUE, chisq = TRUE, expected = TRUE, sresid=TRUE, format="SPSS")
```

## Interpret Results

-   `p < 0.5` indicates significance.

-   Loan term *does* influence loan status.

-   Loans with term of `36 months` are more likley to be fully paid off.

------------------------------------------------------------------------

# Part 2:

## Question Setup

`How has the ability to own a home changed after 2009?`

## Data Wrangling

#### Recode by `Year` for McNemar Chi-Square:

Reformat the `Date` column into a new column, \
Isolate `Year` from `Date` and add as a new column.

```{r}
loans$DateR <- as.Date(paste(loans$Date), "%m/%d/%Y")
DateR <- c(loans$DateR)
loans$DateYear <- format(DateR, format="%Y")
View(loans)
```

```{r}
loans$DateYearR <- NA
loans$DateYearR [loans$DateYear < "2009"] <- 0
loans$DateYearR [loans$DateYear == "2009"] <- 0
loans$DateYearR [loans$DateYear > "2009"] <- 1
```

#### Recode `Own` vs. `Rent`

```{r}
loans$RentvsOwn <- NA
loans$RentvsOwn[loans$home_ownership == "RENT"] <- 0
loans$RentvsOwn[loans$home_ownership == "OWN"] <- 1
```

## Test assumptions and run analysis

```{r}
CrossTable(loans$RentvsOwn, loans$DateYearR, fisher=TRUE, chisq = TRUE, mcnemar = TRUE, expected = TRUE, sresid=TRUE, format="SPSS")
```

#### Check assumption of Expected Frequencies

In order to meet the assumption for the McNemar Chi-Square, you will need to have at least 5 expected per cell. **Has met this assumption.**

## Interpret Results

-   `p < 0.5` indicates significance.

-   There *is a difference* in the ability to own a home after 2009.

## Post Hocs

None of the standardized residuals are over 2. This indicates that the test determined that these results are similar enough that it is not significant.

     2.505% |   11.422%

------------------------------------------------------------------------

# Part 3: 

## Question Setup

The news just ran a story that only 15% of homes are fully paid for in America, and that another 10% have given up on paying it back, so the bank has "charged off" the loan.

`Does it seem likely that the data for this hands-on came from the larger population of America?`

## Data Wrangling

In order to run a goodness of fit Chi-Square, you need the observed values for:\
- home loans that are paid in full\
- home loans that are "charged off"

Get these values by aggregating your data using the `dplyr` `summarize()` function for count (`n()`):

```{r}
loans %>% group_by(loan_status) %>% summarize(count=n())
```

## Run Analysis

Define the observed values as vectors:

```{r}
# ORDER: Charged Off, Current, Fully Paid
observed = c(3282, 502, 18173)
```

Define a vector of expected values. **These expected values must equal 1.** \
From story, "Fully Paid" was `15%` or `.15`, while "Charged Off" was `10%` or `.1`

```{r}
# ORDER: Charged Off, Current, Fully Paid
expected = c(0.1, .75, 0.15)
```

Run the analysis with the function `chisq.test()`, in which you will define the argument of `x=` with your `observed` frequencies, and in which you will define the argument of `p=` with your expected frequencies:

```{r}
chisq.test(x = observed, p = expected)
```

## Interpret Results

**`The p value is less than .05` -\> the observed and expected values differ.**\
*It does NOT seem likely* that this sample came from the larger population of America.

------------------------------------------------------------------------

# `### END`
