---
title: "DS105-04-06 - ANOVAs - One Way Between Subjects - Activity in R"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## Load Libraries

```{r}
library("dplyr")
library("rcompanion")
library("car")
```

------------------------------------------------------------------------

## Load in Data

```{r}
youtube <- read.csv("./assets/YouTubeChannels.csv")

  # View the data
    head(youtube)
```

------------------------------------------------------------------------

## Question Setup

Is there a difference in the **number of views `Video.views`** compared to all the different **grade categories `Grade`**?

`Grade` categories:

```{r}
sort(unique(youtube$Grade))
```

------------------------------------------------------------------------

## Data Wrangling

No data wrangling needed.

------------------------------------------------------------------------

## **Test Assumptions**

### **Assumption: Normality**

Test for the normality of the dependent variable.

```{r}
plotNormalHistogram(youtube$Video.views)
```

#### Transform for normality using `Tukey's Ladder of Power`

```{r}
youtube$Video.viewsTUK <- transformTukey(youtube$Video.views, plotit=FALSE)

plotNormalHistogram(youtube$Video.viewsTUK)
```

::: {style="color: #238823"}
**Successfully transformed, meets assumption of Normality.**
:::

### **Assumption: Homogeneity of Variance**

Using `Bartlett's test` since the data is now transformed close to normal distributed.\
Homogeneity of Variance is met if `p > .05` - *not significant*.

#### **Bartlett's Test**

Use the function `bartlett.test()`, with the argument of the `y data` separated by a tilde, followed by the `x data`. `data=` argument is the name of the dataset.

```{r}
bartlett.test(Video.views ~ Grade, data=youtube)
```

::: {style="color: #D2222D"}
**Fails assumption of Homogeneity of Variance.**

**`p < .05`**, which means that the data fails the assumption of Homogeneity of Variance.
:::

#### **Computing ANOVAs with Unequal Variance (Violated Homogeneity of Variance Assumption)**

```{r}
youtubeANOVA <- lm(Video.viewsTUK ~ Grade, data=youtube)
Anova(youtubeANOVA, Type='II', white.adjust=TRUE)
```

### **Assumption: Sample Size**

*An ANOVA requires a sample size of at least 20 per independent variable.*\
\# of independent variables: 1\
required number of cases: 20.

Number of cases: 5000

```{r}
str(youtube)
```

::: {style="color: #238823"}
**Meets assumption of adequate sample size.**
:::

### **Independence**

There is no statistical test for the assumption of independence.

### **Computing Post Hocs with Unequal Variance (Violated Homogeneity of Variance Assumption)**

```{r}
pairwise.t.test(youtube$Video.viewsTUK, youtube$Grade, p.adjust="bonferroni", pool.sd = FALSE)
```

::: {style="color: #238823"}
**The pairwise comparison indicates there is significant difference.**
:::

------------------------------------------------------------------------

## **Determine Means and Draw Conclusions**

If there is a significant difference after correction, examine the means to interpret results and draw conclusions.

```{r}
youtubeMeans <- youtube %>% group_by(Grade) %>% summarize(Mean = mean(Video.views))

# View tibble sorted by Mean column descending order
youtubeMeans[order(-youtubeMeans$Mean),]
```

::: {style="color: #0187DE"}
# Conclusion:

The number of views for each `Grade` level varies significantly, with higher grades receiving more views and lower grades receiving fewer views.
:::
