---
title: "DS105-05-07 - ANOVAs - Honey Production Practice Hands-On"
output: html_notebook
---

## Load Libraries

```{r}
#install.packages("fastR2")
#install.packages("car")
library("rcompanion")
library("fastR2")
library("car")
library("dplyr")
```

------------------------------------------------------------------------

## Load in Data

```{r}
honey <- read.csv("./assets/honey.csv")

  # View data
    head(honey)
```

------------------------------------------------------------------------

## Question Setup

**Has total honey production (`totalprod`) changed over the years (`year`)?**

------------------------------------------------------------------------

## Data Wrangling

### Filter Data

```{r}
honey$year <- as.character(honey$year)
honey$year <- as.factor(honey$year)

head(honey)
```

------------------------------------------------------------------------

## **Test Assumptions**

### **Assumption: Normality**

You only need to test for the normality of the dependent variable, but you need to do it at both timepoints. So scoot on back to `breakfast2`, which is the dataset you truncated but had not yet reshaped.

```{r}
plotNormalHistogram(honey$totalprod)
```

```{r}
honey$totalprodTUK <- transformTukey(honey$totalprod, plotit=FALSE)

plotNormalHistogram(honey$totalprodTUK)
```

::: {style="color: #238823"}
**Meets assumption of Normality.**
:::

\
**Assumption: Homogeneity of Variance**

```{r}
leveneTest(totalprodTUK ~ year, data=honey)
```

::: {style="color: #238823"}
**Meets assumption of Homogeneity of Variance.** \
**`p > .05`**, which means that the data meets the assumption of Homogeneity of Variance.
:::

### Sample Size

A repeated measures ANOVA requires a sample size of at least 20 per independent variable.

```{r}
str(honey)
```

::: {style="color: #238823"}
**Meets assumption of adequate sample size.**
:::

------------------------------------------------------------------------

## **Analysis**

You will continue to use the `aov()` function, but add some additional arguments to it to make it repeated measures.

```{r}
RManova <- aov(totalprodTUK~year+Error(state), honey)
summary(RManova)
```

```{r}
RManova <- aov(log(totalprod)~year, honey)
summary(RManova)
```

::: {style="color: #0187DE"}
# Conclusion:

***`p`*** **`< .5`**, which indicates ***there is a significant effect*** of year on total honey production.
:::
