---
title: "DS105-08-06 - MANOVA in R Hands-On"
output: html_notebook
---

# Load Libraries

The MANOVA function comes in the base package of R, so the libraries that you will need to load are all related to assumption testing. You will need the following: `mvnormtest` to test for multivariate normality, and `car` to test for homogeneity of variance.

```{r}
# install.packages("mvnormtest")
library("mvnormtest")
library("car")
```

# Load Data

```{r}
heartAttacks <- read.csv("./data/heartAttacks.csv")

  # View data
    head(heartAttacks)
```

# Question Setup

Please answer the following question with this data:

<div>

It is well-known that men are more likely to have heart attacks than women.\
How does gender (`sex`) influence some of the heart attack predictors like resting blood pressure (`trestbps`) and cholesterol (`chol`)?

</div>

**Independent Variable:** Gender ( `sex` )\
**Dependent Variables:** Resting Blood Pressure ( `trestbps` ), Cholesterol ( `chol` )

------------------------------------------------------------------------

# Data Wrangling

To test for multivariate normality,

1.  Create a dataset containing only the *dependent variables*

2.  Ensure that the dependent variables dataset is in a *matrix format*, and

3.  Ensure that the dependent variables are *numeric*

4.  Limit your data to 5,000 rows (the test for normality can only handle 5,000 records)

### Ensure Dependent Variables are Numeric

Check the structure of the data to see what format the dependent variables are in:

```{r}
str(heartAttacks$trestbps)
str(heartAttacks$chol)
```

::: {style="color: #238823"}
**The dependent variables are already integers (numeric format).**
:::

### Subsetting

Keep only the dependent variables:

```{r}
keeps <- c("trestbps", "chol")
heartAttacks1 <- heartAttacks[keeps]
```

Then limit the number of rows to 5000:

```{r}
heartAttacks2 <- heartAttacks1[1:5000,]

  # Confirm the limited dataset
    str(heartAttacks2)
```

### Format data set as a Matrix

```{r}
heartAttacks3 <- as.matrix(heartAttacks2)

  # Confirm structure change of the dataset
    is.matrix(heartAttacks3)
```

Now ready to perform the assumptions test for multivariate normality on `heartAttacks3`.

------------------------------------------------------------------------

# Test Assumptions

### **Sample Size**

Requirement:

-   at least 20 for 1 independent variable

-   in each cell, the number of cases \> number of dependent variables

```{r}
str(heartAttacks)
```

::: {style="color: #238823"}
### Meets assumption: Sample Size

Requirement: 20 samples per Independent Variable = 20, \
at least 2 cases per cell for 2 Dependent Variables. \
**303 observations present in data**.
:::

------------------------------------------------------------------------

### **Multivariate Normality**

To test for multivariate normality, first drop any missing values:

```{r}
heartAttacks4 <- na.omit(heartAttacks3)
```

Test the dataset with `Wilks-Shapiro test`:

```{r}
mshapiro.test(t(heartAttacks4))
```

::: {style="color: #D2222D"}
### Violates assumption: Multivariate Normality

**The *p* value is significant at *p* \< .05**, indicating that the data violates the assumption of multivariate normality.
:::

------------------------------------------------------------------------

### Homogeneity of Variance

Test for homogeneity of variance of both dependent variables with `Levene's Test`.

#### **Dependent Variable: `trestbps`**

```{r}
leveneTest(heartAttacks$trestbps, heartAttacks$sex, data=heartAttacks)
```

::: {style="color: #238823"}
#### Meets assumption: Homogeneity of Variance

**The *p* value is NOT significant at *p* \> .05**, indicating that the data meets the assumption of homogeneity of variance.
:::

#### **Dependent Variable: `chol`**

```{r}
leveneTest(heartAttacks$chol, heartAttacks$sex, data=chol)
```

::: {style="color: #D2222D"}
#### **Violates assumption: Homogeneity of Variance**

**The *p* value is significant at *p* \< .05**, indicating that the data violates the assumption of homogeneity of variance.
:::

------------------------------------------------------------------------

### **Absence of Multicollinearity**

Typically, multicollinearity can be assessed simply by running correlations of your dependent variables with each other. A general rule of thumb is that anything above approximately .7 for correlation (i.e. a strong correlation) indicates the presence of multicollinearity.

Check out the correlation between dependent variables with a simple `cor.test()` function:

```{r}
cor.test(heartAttacks$trestbps, heartAttacks$chol, method="pearson", use="complete.obs")
```

::: {style="color: #238823"}
### Meets assumption: **Absence of Multicollinearity**

**The variables have weak correlation under 0.7 at r = 0.1231742**, indicating that the dependent variables lack multicollinearity.
:::

------------------------------------------------------------------------

# Run Analysis

```{r}
MANOVA <- manova(cbind(trestbps, chol) ~ sex, data = heartAttacks)
summary(MANOVA)
```

::: {style="color: #0187DE"}
### Analysis: 

**There is a significant difference in heart attack predictors and sex of participant.**

**The *p* value is significant at *p* \< .05**, indicating a significant difference in the bound dependent variables and the independent variable.
:::

------------------------------------------------------------------------

# Post Hocs

### ANOVAs as Post Hocs

```{r}
summary.aov(MANOVA, test = "wilks") 
```

::: {style="color: #0187DE"}
### Post Hocs: 

-   [***p*** **value is NOT significant at *p* \> 0.5**, indicating there is a NOT a significant difference in **resting blood pressure** and **sex** of participant. ]{style="color: #D2222D"}

-   ***p*** **value IS significant at *p* \< 0.5**, indicating there IS a significant difference in **cholesterol** and **sex** of participant.
:::

------------------------------------------------------------------------

::: {style="color: #0187DE"}
# Conclusion

> How does gender (`sex`) influence some of the heart attack predictors like resting blood pressure (`trestbps`) and cholesterol (`chol`)?

**Gender affects cholesterol levels significantly,** but does not significantly impact resting blood pressure as a predictor of heart attack.
:::
