---
title: "DS105-06-04 - Mixed Measure ANOVA in R Activity"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

## Load Libraries

```{r}
library("rcompanion")
library("car")
```

------------------------------------------------------------------------

## Load in Data

```{r}
breakfast <- read.csv("./assets/breakfast.csv")

  # View data
    head(breakfast)
```

------------------------------------------------------------------------

## Question Setup

With this data, you will answer the question:

    Did those who ate breakfast in the morning improve their resting metabolic rate from baseline to follow up compared to those who skipped breakfast?  

In order to answer this question, your first x, or independent variable,
will be the `Treatment.Group` of whether they ate breakfast in the
morning or not. Your second x will be time, whether it was baseline or
follow up. Your y, or dependent variable, will be resting metabolic
rate. As with all ANOVAs, the IV will be categorical, and the DV will be
continuous.

------------------------------------------------------------------------

## Data Wrangling

Data wrangling for the mixed measures ANOVA is done exactly the same as
it was for the repeated measures ANOVA.

### Filter Data

```{r}
breakfast1 <- breakfast[1:33,]

head(breakfast1)
```

### Reshape Data

```{r}
keep = c("Participant.Code", "Treatment.Group", "Age..y.", "Sex", "Height..m.", "Baseline.Body.Mass..kg.", "Follow.Up.Body.Mass..kg.")

breakfast2 <- breakfast1[keep]

head(breakfast2)
```

------------------------------------------------------------------------

## Baseline Data

```{r}
breakfast3 <- breakfast2[,1:5]
breakfast3$repdat <- breakfast2$Baseline.Body.Mass..kg.
breakfast3$contrasts <- "T1"
head(breakfast3)
```

## Follow-up Data

```{r}
breakfast4 <- breakfast2[,1:5]
breakfast4$repdat <- breakfast2$Follow.Up.Body.Mass..kg.
breakfast4$contrasts <- "T2"
head(breakfast4)
```

Once you have both of those, then you need to `rbind()` them back
together into one whole dataset:

```{r}
breakfast5 <- rbind(breakfast3, breakfast4)
```

Now you are all prepared to run a repeated measures ANOVA, data shaping
wise.

------------------------------------------------------------------------

## **Testing Assumptions**

The assumptions for a mixed measure ANOVA are the same as the ones you
learned for a repeated measures ANOVA. The only thing that differs is
the sample size, because you now have two IVs.

------------------------------------------------------------------------

### **Sample Size**

A mixed measures ANOVA requires a sample size of at least 20 per
independent variable or time factor. In this case, you only have one
independent variable, and you also have a factor of time. So, you need
40 cases. You are a few cases short of this requirement, clocking in at
only *n* = 33, but for learning purposes, you will proceed. However,
typically if you did not have a large enough sample size, you would
either want to simplify your model (remove either the IV or the time
variable), choose a different analysis, or run a procedure called
bootstrapping which would re-sample your data until you had a larger
*n*.

------------------------------------------------------------------------

## Analysis

You will continue to use the `aov()` function, but add some additional
arguments to it to make it mixed measures.

```{r}
RManova1 <- aov(repdat~(Treatment.Group*contrasts)+Error(Participant.Code/(contrasts)), breakfast5)
summary(RManova1)
```

So what's happening here is that you are calling the `aov()` function on
the interaction between all of your data factors. First, you are saying
that you want to see the repeated data of resting metabolic rate by your
time factor (from baseline to follow-up). That's this part:
`repdat~(Treatment.Group*contrasts`. Next, you are adding in your error
term, which is specified in this model by the command `Error()`. In the
error term, you are placing your subject identifier (which matches the
pre and the post data together), and you also note that it needs to be
done for both time factor groups as well. That's what this part of the
code is doing: `+Error(Participant.Code/(contrasts)`. Finish it all off
by specifying the dataset at the end and you are good to go. Call
`summary()` on the results:

------------------------------------------------------------------------

## Post Hocs

After finding such a load of bupkis above, so no need to worry about
post hocs here!

------------------------------------------------------------------------

# `### END` 
