---
title: "DS105-02-11 - Python and R Hands-On - Seattle Parks and Rec"
output:
  html_document:
    df_print: paged
---

### Load Packages

```{r}
library("rcompanion")
library("IDPmisc")
```

### Load Data

```{r}
parksrec <- read.csv("./assets/Seattle_ParksnRec.csv")
```

### View Data

```{r}
View(parksrec)
```

------------------------------------------------------------------------

### Finding Column Names

```{r}
colnames(parksrec)
```

------------------------------------------------------------------------

# Variable: `X..of.trips.Fall`

::: {style="color: blue"}
positively skewed -\> Transformed with Tukey's Ladder of Power
:::

```{r}
# PLOT NORMAL HISTORGRAM FOR VARIABLE
plotNormalHistogram(parksrec$"X..of.trips.Fall")
```

```{r}
# TRANSFORM:
parksrec$"X..of.trips.FallTUK" <- transformTukey(parksrec$"X..of.trips.Fall", plotit=FALSE)

# RUN HISTOGRAM AGAIN
plotNormalHistogram(parksrec$"X..of.trips.FallTUK")
```

------------------------------------------------------------------------

# Variable: `X..of.participants.Fall`

::: {style="color: blue"}
positively skewed -\> Transformed with Tukey's Ladder of Power
:::

```{r}
# PLOT NORMAL HISTORGRAM FOR VARIABLE
plotNormalHistogram(parksrec$"X..of.participants.Fall")
```

```{r}
# TRANSFORM:
parksrec$"X..of.participants.FallTUK" <- transformTukey(parksrec$"X..of.participants.Fall", plotit=FALSE)

# RUN HISTOGRAM AGAIN
plotNormalHistogram(parksrec$"X..of.participants.FallTUK")
```

------------------------------------------------------------------------

# Variable: `X..of.trips.per.year`

::: {style="color: blue"}
positively skewed -\> Transformed with Tukey's Ladder of Power
:::

```{r}
# PLOT NORMAL HISTORGRAM FOR VARIABLE
plotNormalHistogram(parksrec$"X..of.trips.per.year")
```

```{r}
# TRANSFORM:
parksrec$"X..of.trips.per.yearTUK" <- transformTukey(parksrec$"X..of.trips.per.year", plotit=FALSE)

# RUN HISTOGRAM AGAIN
plotNormalHistogram(parksrec$"X..of.trips.per.yearTUK")
```

------------------------------------------------------------------------

# Variable: `X..of.participants.per.Year`

::: {style="color: blue"}
positively skewed -\> Transformed with Tukey's Ladder of Power
:::

```{r}
# PLOT NORMAL HISTORGRAM FOR VARIABLE
plotNormalHistogram(parksrec$"X..participants.per.year")
```

```{r}
# TRANSFORM:
parksrec$"X..participants.per.yearTUK" <- transformTukey(parksrec$"X..participants.per.year", plotit=FALSE)

# RUN HISTOGRAM AGAIN
plotNormalHistogram(parksrec$"X..participants.per.yearTUK")
```

------------------------------------------------------------------------

# Variable: `increase.decrease.of.prior.year`

::: {style="color: blue"}
positively skewed, but close to normal -\> Transformed with Tukey's Ladder of Power
:::

```{r}
# PLOT NORMAL HISTORGRAM FOR VARIABLE
plotNormalHistogram(parksrec$"increase.decrease.of.prior.year")
```

```{r}
# TRANSFORM:
parksrec$"increase.decrease.of.prior.yearTUK" <- transformTukey(parksrec$"increase.decrease.of.prior.year", plotit=FALSE)

# RUN HISTOGRAM AGAIN
plotNormalHistogram(parksrec$"increase.decrease.of.prior.yearTUK")
```

------------------------------------------------------------------------

# Variable: `Average...people.per.trip`

::: {style="color: blue"}
just barely positively skewed -\> Transformed with Tukey's Ladder of Power
:::

```{r}
# PLOT NORMAL HISTORGRAM FOR VARIABLE
plotNormalHistogram(parksrec$"Average...people.per.trip")
```

```{r}
# TRANSFORM:
parksrec$"Average...people.per.tripTUK" <- transformTukey(parksrec$"Average...people.per.trip", plotit=FALSE)

# RUN HISTOGRAM AGAIN
plotNormalHistogram(parksrec$"Average...people.per.tripTUK")
```

# `### END`
