---
title: "DS105-06-07 - Lesson 6 Hands-On - in R"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

## Install Packages

```{r}
install.packages("rcompanion")
install.packages("car")
install.packages("IDPmisc")
install.packages("dplyr")
install.packages("tidyverse")
```

## Load Libraries

```{r}
library("rcompanion")
library("car")
library("IDPmisc")
library("dplyr")
library(tidyverse)
```

------------------------------------------------------------------------

## Load in Data

```{r}
suicide <- read.csv("./assets/suicide.csv")

  # View data
    head(suicide)
```

```{r}
str(suicide)
```

------------------------------------------------------------------------

## Question Setup

You will determine whether suicide rates (`suicides/100k pop`) has
changed over the years (`year`), and see if the `generation` has any
influence. To do so, you will be using a mixed measures ANOVA, since
there is both a repeated time element and a between subjects element.

------------------------------------------------------------------------

## Data Wrangling

Data wrangling for the mixed measures ANOVA is done exactly the same as
it was for the repeated measures ANOVA.

------------------------------------------------------------------------

## Baseline Data

```{r}
breakfast3 <- breakfast2[,1:5]
breakfast3$repdat <- breakfast2$Baseline.Body.Mass..kg.
breakfast3$contrasts <- "T1"
head(breakfast3)
```

## Follow-up Data

```{r}
breakfast4 <- breakfast2[,1:5]
breakfast4$repdat <- breakfast2$Follow.Up.Body.Mass..kg.
breakfast4$contrasts <- "T2"
head(breakfast4)
```

Once you have both of those, then you need to `rbind()` them back
together into one whole dataset:

```{r}
breakfast5 <- rbind(breakfast3, breakfast4)
```

------------------------------------------------------------------------

## **Testing Assumptions**

------------------------------------------------------------------------

### Assumption: Normality

```{r}
plotNormalHistogram(suicide$suicides.100k.pop)
```

### Transformation: Square Root

```{r}
suicide$suicides.100k.popSQRT <- sqrt(suicide$suicides.100k.pop)

plotNormalHistogram(suicide$suicides.100k.popSQRT)
```

### Transformation: Logarithmic

```{r}
suicide$suicides.100k.popLOG <- log(suicide$suicides.100k.pop)

# Remove `null` values after getting an error graphing LOG transformation.
suicide2 <- NaRV.omit(suicide)
plotNormalHistogram(suicide2$suicides.100k.popLOG)
```

::: {style="color: #238823"}
**After Transformation, meets assumption of Normality.**
:::

------------------------------------------------------------------------

### Assumption: Homogeneity of Variance

```{r}
leveneTest(suicides.100k.popLOG ~ generation, data=suicide2)
```

::: {style="color: #D2222D"}
**Fails assumption of Homogeneity of Variance.**

**`p < .05`**, which means that the data fails the assumption of
Homogeneity of Variance.
:::

------------------------------------------------------------------------

### **Assumption: Sample Size**

A mixed measures ANOVA requires a sample size of at least 20 per
independent variable or time factor.

We have 1 IV and 1 factor of time = 2 x 20 = need 40 minimum. We have
`27820` observations.

```{r}
str(suicide)
```

::: {style="color: #238823"}
**Meets the assumption of Sample Size.**
:::

------------------------------------------------------------------------

## Analysis

You will continue to use the `aov()` function, but add some additional
arguments to it to make it mixed measures.

```{r}
RManova <- aov(suicides.100k.popLOG~(generation*year)+Error(country/(year)), suicide2)
summary(RManova)
```

::: {style="color: #0187DE"}
# Conclusion:

***`p`*** **`< .5`** which indicates there *is a significant effect* of
**`time`** and **`generation`** on suicides rates.
:::

------------------------------------------------------------------------

# `### END`
