---
title: "DS105-05-02 - ANOVAs - Repeated Measures - in R"
output: html_notebook
---

## Load Libraries

```{r}
#install.packages("fastR2")
#install.packages("car")
library("rcompanion")
library("fastR2")
library("car")
```

------------------------------------------------------------------------

## Load in Data

```{r}
breakfast <- read.csv("./assets/breakfast.csv")

  # View data
    head(breakfast)
```

------------------------------------------------------------------------

## Question Setup

**Were there weight changes from baseline to follow?**

------------------------------------------------------------------------

## Data Wrangling

### Filter Data

```{r}
breakfast1 <- breakfast[1:33,]

head(breakfast1)
```

### Reshape Data

```{r}
keep = c("Participant.Code", "Treatment.Group", "Age..y.", "Sex", "Height..m.", "Baseline.Body.Mass..kg.", "Follow.Up.Body.Mass..kg.")

breakfast2 <- breakfast1[keep]

head(breakfast2)
```

------------------------------------------------------------------------

## Baseline Data

```{r}
breakfast3 <- breakfast2[,1:5]
breakfast3$repdat <- breakfast2$Baseline.Body.Mass..kg.
breakfast3$contrasts <- "T1"
head(breakfast3)
```

## Follow-up Data

```{r}
breakfast4 <- breakfast2[,1:5]
breakfast4$repdat <- breakfast2$Follow.Up.Body.Mass..kg.
breakfast4$contrasts <- "T2"
head(breakfast4)
```

Once you have both of those, then you need to `rbind()` them back together into one whole dataset:

```{r}
breakfast5 <- rbind(breakfast3, breakfast4)
```

Now you are all prepared to run a repeated measures ANOVA, data shaping wise.

\

------------------------------------------------------------------------

## **Test Assumptions**

### **Assumption: Normality**

You only need to test for the normality of the dependent variable, but you need to do it at both timepoints. So scoot on back to `breakfast2`, which is the dataset you truncated but had not yet reshaped.

```{r}
plotNormalHistogram(breakfast2$Baseline.Body.Mass..kg.)
```

Try the follow-up data:

```{r}
plotNormalHistogram(breakfast2$Follow.Up.Body.Mass..kg.)
```

::: {style="color: #238823"}
**Meets assumption of Normality.**
:::

\
**Assumption: Homogeneity of Variance**

```{r}
leveneTest(repdat ~ Treatment.Group*contrasts, data=breakfast5)
```

::: {style="color: #238823"}
**Meets assumption of Homogeneity of Variance.** \
**`p > .05`**, which means that the data meets the assumption of Homogeneity of Variance.
:::

### Sample Size

A repeated measures ANOVA requires a sample size of at least 20 per independent variable.

```{r}
str(breakfast5)
```

::: {style="color: #238823"}
**Meets assumption of adequate sample size.**
:::

### Sphericity

The only way to test for sphericity in R is to take a multivariate approach and make it work for an ANOVA. At this time, that is a bit too complex, but it may be covered later.

------------------------------------------------------------------------

## **Analysis**

You will continue to use the `aov()` function, but add some additional arguments to it to make it repeated measures.

```{r}
RManova <- aov(repdat~contrasts+Error(Participant.Code), breakfast5)
summary(RManova)
```

So what's happening here is that you are calling the `aov()` function on your repeated data of metabolic rate by your timepoint, and then adding in an error term, which is what makes this a repeated measure - you are telling it that it should be looking within each participant, which is what ths part of the code does: `Error(Participant.Code)`. Finish it all off by specifying the dataset at the end and you are good to go. Call `summary()` on the results:

\
Under the `Error:Within` table (since this is a within subjects ANOVA, after all), the you will find your *F* value and the associated *p* value. Looks like there is not a significant effect of time on resting metabolic rate.

::: {style="color: #0187DE"}
# Conclusion:

***`p`*** **`> .5`**, which indicates there is *not a significant effect* of time on the participants' body mass.
:::
