---
title: "DS106-04-08 - Lesson 4 Practice Hands-On - Stepwise Regression - Part 2"
output:
  html_document:
    df_print: paged
---

```{r}
stepwiseRegression = read.csv("data/stepwiseRegression.csv")

head(stepwiseRegression)
```

```{r}
# BACKWARD ELIMINATION

# Create a linear model of all predictor variables.
FitAll = lm(Y ~ ., data = stepwiseRegression)

# Summary of model
summary(FitAll)

step(FitAll, direction = 'backward')

```

```{r}
# FORWARD SELECTION

# Create a linear model of all predictor variables.
FitStart = lm(Y ~ 1, data = stepwiseRegression)

# Summary of model
summary(FitStart)

step(FitStart, direction = 'forward', scope = (formula(FitAll)))

```

```{r}
# HYBRID STEPWISE

# Create model
step(FitStart, direction = 'both', scope = (formula(FitAll)))

```

# Questions:

1.  Which model was the best for each method?
    1.  Backward Elimination:

            formula = Y ~ X2 + X4 + X6 + X10 + X11 + X12

    2.  Forward Selection:

            formula = Y ~ X6 + X4 + X12 + X10 + X2 + X11

    3.  Hybrid Stepwise:

            formula = Y ~ X6 + X4 + X12 + X10 + X2 + X11

2.  How do the final models from each method compare to each other?\
    They each include: `X2`, `X4`, `X6`, `X10`, `X11`, and `X12`.\
    However, the order is the *same* for Forward Selection and Hybrid Stepwise, \
    while the order is drastically different for Backward Elimination.
