---
title: "DS106-04-08 - Lesson 4 Practice Hands-On - Stepwise Regression - Part 1"
output:
  html_document:
    df_print: paged
---

`{# Part 1: Backwards Elimination}`

```{r}
iq = read.csv("data/IQ.csv")
```

```{r}
head(iq)
```

```{r}
# Create a linear model of all 10 predictor variables.
FitAll = lm(IQ ~ ., data = iq)
```

```{r}
# Summary of model
summary(FitAll)
```

```{r}
step(FitAll, direction = 'backward')
```

```{r}
# Create new model with just the un-eliminated variables:
FitSome = lm(formula = IQ ~ Test1 + Test2 + Test4, data = iq)

# Summary of model
summary(FitSome)
```

# Questions:

1.  Which model is the best? Why?\
    Best model is:

        IQ ~ Test1 + Test2 + Test4

2.  From the best model, what is the adjusted R\^2 value and what does it mean?

    1.  There is a `multiple R2 = 0.3839`. The model explains 38.39% of the variation the `IQ` variable.

    2.  There is an `adjusted R2 = 0.2158`. The adjustment is a modification that is supposed to take into account the number of terms in the model.

3.  From the best model, how does each variable influence IQ?

    1.  `Test1` affects `IQ` by `-1.9650` -\> as `Test1` goes down, `IQ` goes up.

    2.  `Test2` affects `IQ` by `-1.6485` -\> as `Test2` goes down, `IQ` goes up.

    3.  `Test4` affects `IQ` by `3.7890` -\> as `Test4` goes up, `IQ` also goes up.
