---
title: "DS105-04-13 - Lesson 4 Hands-On - Analyzing Sales - in R"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## Load Libraries

```{r}
library("dplyr")
library("rcompanion")
library("car")
library("tidyverse")
```

------------------------------------------------------------------------

## Load in Data

```{r}
avocados <- read.csv("./assets/avocados.csv")

  # View data
    head(avocados)
```

------------------------------------------------------------------------

## Question Setup

**Does the average price of avocados differ between Albany, Houston, and Seattle?**

```{r}
# Find column names
colnames(avocados)
```

------------------------------------------------------------------------

## Data Wrangling

### Filter Data

```{r}
price = na.omit(avocados %>% filter(region %in% c("Albany", "Houston", "Seattle")))

  # Confirm data change
    unique(price$region)
```

### Convert Dependent Variable `price` to numberic

```{r}
price$AveragePrice = as.numeric(price$AveragePrice)

  # Confirm data change
    str(price)
```

------------------------------------------------------------------------

## **Test Assumptions**

### **Assumption: Normality**

Test for the normality of the dependent variable.

```{r}
plotNormalHistogram(price$AveragePrice)
```

#### Transform for normality using `Tukey's Ladder of Power`

```{r}
price$AveragePriceTUK <- transformTukey(price$AveragePrice, plotit=FALSE)

plotNormalHistogram(price$AveragePriceTUK)
```

::: {style="color: #238823"}
**Successfully transformed, meets assumption of Normality.**
:::

### **Assumption: Homogeneity of Variance**

Using `Bartlett's test` since the data is now transformed close to normal distributed.\
Homogeneity of Variance is met if `p > .05` - *not significant*.

#### **Bartlett's Test**

Use the function `bartlett.test()`, with the argument of the `y data` separated by a tilde, followed by the `x data`. `data=` argument is the name of the dataset.

```{r}
bartlett.test(AveragePriceTUK ~ region, data=price)
```

::: {style="color: #D2222D"}
**Fails assumption of Homogeneity of Variance.**

**`p < .05`**, which means that the data fails the assumption of Homogeneity of Variance.
:::

#### **Computing ANOVAs with Unequal Variance (Violated Homogeneity of Variance Assumption)**

```{r}
priceANOVA <- lm(AveragePriceTUK ~ region, data=price)
Anova(priceANOVA, Type='II', white.adjust=TRUE)
```

### **Assumption: Sample Size**

*An ANOVA requires a sample size of at least 20 per independent variable.*\
\# of independent variables: 1\
required number of cases: 20

Number of cases: 1014

```{r}
str(price)
```

::: {style="color: #238823"}
**Meets assumption of adequate sample size.**
:::

### **Independence**

There is no statistical test for the assumption of independence.

### **Computing Post Hocs with Unequal Variance (Violated Homogeneity of Variance Assumption)**

```{r}
pairwise.t.test(price$AveragePriceTUK, price$region, p.adjust="bonferroni", pool.sd = FALSE)
```

::: {style="color: #238823"}
**The pairwise comparison indicates there is significant difference.**
:::

------------------------------------------------------------------------

## **Determine Means and Draw Conclusions**

If there is a significant difference after correction, examine the means to interpret results and draw conclusions.

```{r}
priceMeans <- price %>% group_by(region) %>% summarize(Mean = mean(AveragePrice))

# View tibble sorted by Mean column descending order
priceMeans[order(-priceMeans$Mean),]
```

::: {style="color: #0187DE"}
# Conclusion:

The `Average Price` of avocados differs between the three regions, with Albany being the highest at \$1.56, Seattle at \$1.44, and Houston the lowest at \$1.05.
:::
